---
title: ""
format:
  html:
    toc: true
    toc-depth: 4
    toc-expand: true
    code-link: false
    css: styles.css
    page-layout: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```



## Data Transformation Techniques  
Data transformations can help correct skewness and stabilize variance. MVN supports simple transforms (log, square root, square) via the `transform` argument, as well as optimal Box–Cox transformation.

---

## Example Data and Baseline Assessment

```{r}
# Load package and example data
library(MVN)
df <- iris[1:50, 1:4]

# Baseline Henze–Zirkler test without transformation
result <- mvn(data = df, mvn_test = "hz")
summary(result, select = "mvn")
```

The untransformed data are marginally on the threshold of normality (p = 0.05), indicating potential departure from multivariate normality.


## 1. Basic Transformations

Use the `transform` parameter in `mvn()` to apply:

- **"log"** (natural log)
- **"sqrt"** (square root)
- **"square"** (power of 2)

```{r}
# Log-transform
res_log <- mvn(data = df, mvn_test = "hz", transform = "log")
summary(res_log, select = "mvn")
```
The log transformation improves normality (p = 0.379), suggesting data are now consistent with a multivariate normal distribution.

```{r}
# Square-root transform
res_sqrt <- mvn(data = df, mvn_test = "hz", transform = "sqrt")
summary(res_sqrt, select = "mvn")
```

The square-root transformation also corrects skewness, producing p = 0.253 and acceptance of normality.


```{r}
# Square transform
res_sq <- mvn(data = df, mvn_test = "hz", transform = "square")
summary(res_sq, select = "mvn")
```

Squaring amplifies deviations and results in clear non-normality (p < 0.001), so this transform is not recommended.

---

## 2. Box–Cox Transformation


Enable Box–Cox with optimal lambda estimation using:

```{r}
res_bc_sum <- mvn(
  data              = df,
  mvn_test          = "hz",
  box_cox_transform = TRUE
)
summary(res_bc_sum, select = "mvn")
```

The Box–Cox transformation yields p = 0.332, indicating a good fit to multivariate normality with data-driven lambda values applied.

Here are the optimal lambda values used for transformation:

```{r}
res_bc_sum$box_cox_lambda
```

---

## References

Korkmaz S, Goksuluk D, Zararsiz G. MVN: An R Package for Assessing Multivariate Normality. *The R Journal*. 2014;6(2):151–162. URL: https://journal.r-project.org/archive/2014-2/korkmaz-goksuluk-zararsiz.pdf
