---
title: ""
format:
  html:
    toc: true
    toc-depth: 4
    toc-expand: true
    code-link: false
    css: styles.css
    page-layout: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
library(MVN)

# Example data: two variables from iris (Part 1 & 2)
df <- iris[1:50, 1:2]
```

## Multivariate Outlier Detection  

Before interpreting your multivariate normality tests, it’s important to check for and understand any influential outliers. In this section, we’ll:

1. **Detect** multivariate outliers using robust Mahalanobis distances.  
2. **Summarize** flagged observations via the summary method.  
3. **Visualize** outliers in Q–Q and scatter plots.

---

### Example Data {.unnumbered .unlisted}

```{r}
# Load the package:
library(MVN)
```

We’ll use two numeric variables from the built-in `iris` dataset:

```{r}
df <- iris[1:50, 1:2]
head(df)
```

## 1. Detecting Outliers

The `mvn()` function can automatically flag multivariate outliers using methods such as the adjusted quantile approach (`"adj"`) or a fixed quantile cutoff. Specify via `multivariate_outlier_method`:

```{r}
out_res <- mvn(
  data = df,
  mvn_test = "hz",
  multivariate_outlier_method = "quan"
)
```
This computes robust Mahalanobis distances and flags observations above the chi-square cutoff at the specified `alpha` (default 0.05).

---

## 2. Summarizing Outliers

Use the `summary()` function with `select = "outliers"` to list flagged observations:

```{r}
summary(out_res, select = "outliers")
```
The output shows each outlier’s observation index and Mahalanobis distance, helping you decide whether to inspect or remove these points.

---

## 3. Visualizing Outliers

```{r}
plot(out_res, diagnostic = "outlier")
```
This Q–Q plot highlights points deviating from the theoretical chi-square line.

---

## References

Korkmaz S, Goksuluk D, Zararsiz G. MVN: An R Package for Assessing Multivariate Normality. *The R Journal*. 2014;6(2):151–162. URL: https://journal.r-project.org/archive/2014-2/korkmaz-goksuluk-zararsiz.pdf
