---
title: ""
format:
  html:
    toc: true
    toc-depth: 4
    toc-expand: true
    code-link: false
    css: styles.css
    page-layout: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

## Subset Analysis  
Subset analysis lets you assess multivariate normality separately in each level of a factor. This is useful when data structure or experimental design requires group-wise validation.

---


## Example Data

```{r}
library(MVN)

# Remove the 4th column, keep Species as grouping
iris_df <- iris[-4]
head(iris_df)
```

## 1. Running MVN by Group

Specify the `subset` argument in `mvn()`:

```{r}
# Henze–Zirkler test by species
subset_res <- mvn(
  data       = iris_df,
  subset     = "Species",
  mvn_test   = "hz"
)
```

## 2. Viewing Group-Specific Results

Extract multivariate normality for each group:

```{r}
summary(subset_res, select = "mvn")
```


All species groups exhibit multivariate normality (p > 0.05). Group-wise analysis ensures that assumptions hold within each category.

---

## 3. Group-Wise Diagnostics

You can also generate diagnostic plots for each subset by `subset_res` object to `plot()`:

```{r}
# Mahalanobis Q–Q plots for each species
plot(
  subset_res,
  diagnostic = "multivariate",
  type       = "qq"
)
```


---

## References

Korkmaz S, Goksuluk D, Zararsiz G. MVN: An R Package for Assessing Multivariate Normality. *The R Journal*. 2014;6(2):151–162. URL: https://journal.r-project.org/archive/2014-2/korkmaz-goksuluk-zararsiz.pdf
